### 2013-04-28 _posts/2013-04-28-cross-referencing-reviewer-replies-in-latex.md

---
layout: post
title: Cross-referencing Revision Line Numbers in LaTeX
published: true
---

One of the more tedious elements of the scientific publishing process is cross-referencing all your revisions in a reply letter to the editor. Typically you make changes in your manuscript and then refer to those changes in the reply letter by page and line number.

For example, you might make a change to your manuscript and then in your reply letter you refer to it:

    We have revised the text. Please see page 3, line 23.

There are many strategies for dealing with this, but most of them require finding all your changes at the end and adding the page numbers and line numbers in the reply letter. This is fine if you only have to do it once, but inevitably something gets changed in the manuscript and all the cross-references need to be redone. Even worse, in working with Word documents that have track changes enabled I've experienced the line numbers changing after all changes are accepted.

I've settled on a fairly simple solution in LaTeX which requires minimal markup and automatically cross-references my changes. While this example deals with the case where the line numbers restart on each page, the solution is even simpler if you have continuous line numbers.

This solution assumes that your reply document is being written at the end of your manuscript. This isn't the only option, but it makes cross-referencing the simplest. At the end, before submitting the manuscript and reply document, I split the PDF in two.

First of all, you'll need to set up page-wise line numbers. Add the `lineno` package to the preamble:

    \usepackage[pagewise]{lineno}

And wherever you want to start the line numbers in your document add:

    \linenumbers

Now we'll set up two new commands in the preamble. First, create a new command to label your changes:

    \newcommand{\r}[1]{\label{#1}\linelabel{#1}}

Then, create a new command to cross-reference your changes:

    \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}

Now, say you wanted to revise a sentence to deal with Reviewer comment A1. You can just add something like this within the manuscript:

    \r{A1} A sentenced we have revised...

In the reply document, you can reference this change like this::

    We have revised the text. Please see \lr{A1}.

Now, when you compile your document, this will get converted to:

    We have revised the text. Please see page 3, line 23.

And as you modify your manuscript the page and line numbers will update.

### 2013-06-20 cv-body.tex
Update CV

\begin{description}
\item[2011--13]
Ph.D.\ Candidate, Simon Fraser University, Burnaby, Canada Supervisors:
Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper\\Canadian Fulbright
Scholar to the University of Washington in 2012--13
\item[2010]
M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\Thesis title:
Trends, drivers, and ecosystem effects of expanding global invertebrate
fisheries. Supervisor: Dr.\ Heike K. Lotze.
(\href{papers/Anderson_MSc_2010.pdf}{PDF})\\Canadian
\href{http://goo.gl/nA1zE}{Governor General's Academic Gold Medal} for
the top-ranked Master's Natural Sciences and Engineering thesis at
Dalhousie University in 2010
\item[2004--07]
B.Sc.\ (Hons.) Environmental Science, Dalhousie University, Halifax,
Canada\\Thesis title: How sustainable are emerging low-trophic level
fisheries on the Scotian Shelf? Supervisors: Dr.\ Heike K. Lotze and
Dr.\ Ransom A. Myers. (\href{papers/Anderson_HSc_2006.pdf}{PDF})\\First
Class Honours \& Environmental Programmes Honour Society Medal
\item[2001--03]
B.Sc.\ Candidate (Hons.) Kinesiology, University of Waterloo, Waterloo,
Canada Transferred to Dalhousie University in 2004.
\end{description}

@@

\begin{description}
\item[2012]
Phillis, C.C.*, S.M. O'Regan*, S.J. Green*, J.E.B. Bruce*, \textbf{S.C.
Anderson}, J.N. Linton, Earth2Ocean Research Derby, B. Favaro. Multiple
pathways to conservation success. Conservation Letters. Early view
online. \url{http://dx.doi.org/10.1111/j.1755-263X.2012.00294.x}.
(*\emph{Authors contributed equally, listed in reverse alphabetical
order})
\item[2012]
Harnik, P.G., H.K. Lotze, \textbf{S.C. Anderson}, Z.V. Finkel, S.
Finnegan, D.R. Lindberg, L.H. Liow, R. Lockwood, C.R. McClain, J.L.
McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson, D.P. Tittensor.
Extinctions in ancient and modern seas. Trends in Ecology and Evolution.
27(11): 608--617. \url{http://dx.doi.org/10.1016/j.tree.2012.07.010}.
\item[2012]
\textbf{Anderson, S.C.}, T.A. Branch, D. Ricard, H.K. Lotze. Assessing
global marine fishery status with a revised dynamic catch-based method
and stock-assessment reference points. ICES Journal of Marine Science.
69(8): 1491--1500. \url{http://dx.doi.org/10.1093/icesjms/fss105}.
(\href{http://icesjms.oxfordjournals.org/cgi/reprint/fss105?ijkey=pFRzMGEDycomI3X\&keytype=ref}{PDF})
\item[2011]
\textbf{Anderson, S.C.}, B.G. Farmer, F. Ferretti, A.L.S. Houde, and
J.A. Hutchings. Correlates of vertebrate extinction risk in Canada.
BioScience. 61(7): 538--549.
\url{http://dx.doi.org/10.1525/bio.2011.61.7.8}.
(\href{papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF})
\item[2011]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
Serial exploitation of global sea cucumber fisheries. Fish and
Fisheries. 12(3): 317-339.
\url{http://dx.doi.org/10.1111/j.1467-2979.2010.00397.x}.
(\href{papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF})
\href{http://www.sciencemag.org/content/331/6014/129.1.full}{Featured in
Science} on January 14 2011.
\item[2011]
Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down interactions
and temperature control of snow crab abundance in the northwest
Atlantic. Marine Ecology Progress Series. 429: 169--183.
\url{http://dx.doi.org/10.3354/meps09081}.
\item[2011]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze. Rapid
global expansion of invertebrate fisheries: trends, drivers, and
ecosystem effects. PLoS ONE. 6(3): e14735.
\url{http://dx.doi.org/10.1371/journal.pone.0014735}
\item[2008]
\textbf{Anderson, S.C.}, H.K. Lotze, and N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic-level fisheries in Atlantic
Canada. Can. J. Fish. Aquat. Sci. 65(12): 2553--2571.
\url{http://dx.doi.org/10.1139/F08-156}.
(\href{papers/Anderson_etal_2008_knowledge.pdf}{PDF}) Eighth most-read
CJFAS article in 2008.
\end{description}

\section{Submitted manuscripts}

\begin{description}
\item[2013]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Ecological prophets:
Quantifying metapopulation portfolio effects. Submitted with revisions
to Methods in Ecology and Evolution.
\item[2013]
O'Regan, S.M., W.J. Palen, \textbf{S.C. Anderson}. Climate warming
mediates negative impacts of rapid pond drying for three amphibian
species. Submitted to Ecology.
\end{description}

>>>>

\begin{description}
\item[2011--13]
Ph.D.\ Candidate, Simon Fraser University, Burnaby, Canada Supervisors:
Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper Canadian Fulbright
Scholar to the University of Washington in 2012--13
\item[2010]
M.Sc.\ Biology, Dalhousie University, Halifax, Canada Thesis title:
Trends, drivers, and ecosystem effects of expanding global invertebrate
fisheries. Supervisor: Dr.\ Heike K. Lotze.
(\href{papers/Anderson_MSc_2010.pdf}{PDF}) Canadian
\href{http://goo.gl/nA1zE}{Governor General's Academic Gold Medal} for
the top-ranked Master's Natural Sciences and Engineering thesis at
Dalhousie University in 2010
\item[2004--07]
B.Sc.\ (Hons.) Environmental Science, Dalhousie University, Halifax,
Canada Thesis title: How sustainable are emerging low-trophic level
fisheries on the Scotian Shelf? Supervisors: Dr.\ Heike K. Lotze and
Dr.\ Ransom A. Myers. (\href{papers/Anderson_HSc_2006.pdf}{PDF}) First
Class Honours \& Environmental Programmes Honour Society Medal
\item[2001--03]
B.Sc.\ Candidate (Hons.) Kinesiology, University of Waterloo, Waterloo,
Canada Transferred to Dalhousie University in 2004.
\end{description}

@@

\begin{description}
\item[2012]
Phillis, C.C.*, S.M. O'Regan*, S.J. Green*, J.E.B. Bruce*, \textbf{S.C.
Anderson}, J.N. Linton, Earth2Ocean Research Derby, B. Favaro. Multiple
pathways to conservation success. Conservation Letters. Early view
online. \url{http://dx.doi.org/10.1111/j.1755-263X.2012.00294.x}.
(*\emph{Authors contributed equally, listed in reverse alphabetical
order})
\item[2012]
Harnik, P.G., H.K. Lotze, \textbf{S.C. Anderson}, Z.V. Finkel, S.
Finnegan, D.R. Lindberg, L.H. Liow, R. Lockwood, C.R. McClain, J.L.
McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson, D.P. Tittensor.
Extinctions in ancient and modern seas. Trends in Ecology and Evolution.
27(11): 608--617. \url{http://dx.doi.org/10.1016/j.tree.2012.07.010}.
\item[2012]
\textbf{Anderson, S.C.}, T.A. Branch, D. Ricard, H.K. Lotze. Assessing
global marine fishery status with a revised dynamic catch-based method
and stock-assessment reference points. ICES Journal of Marine Science.
69(8): 1491--1500. \url{http://dx.doi.org/10.1093/icesjms/fss105}.
(\href{http://icesjms.oxfordjournals.org/cgi/reprint/fss105?ijkey=pFRzMGEDycomI3X\&keytype=ref}{PDF})
\item[2011]
\textbf{Anderson, S.C.}, B.G. Farmer, F. Ferretti, A.L.S. Houde, and
J.A. Hutchings. Correlates of vertebrate extinction risk in Canada.
BioScience. 61(7): 538--549.
\url{http://dx.doi.org/10.1525/bio.2011.61.7.8}.
(\href{papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF})
\item[2011]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
Serial exploitation of global sea cucumber fisheries. Fish and
Fisheries. 12(3): 317-339.
\url{http://dx.doi.org/10.1111/j.1467-2979.2010.00397.x}.
(\href{papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF})
\href{http://www.sciencemag.org/content/331/6014/129.1.full}{Featured in
Science} on January 14 2011.
\item[2011]
Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down interactions
and temperature control of snow crab abundance in the northwest
Atlantic. Marine Ecology Progress Series. 429: 169--183.
\url{http://dx.doi.org/10.3354/meps09081}.
\item[2011]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze. Rapid
global expansion of invertebrate fisheries: trends, drivers, and
ecosystem effects. PLOS ONE. 6(3): e14735.
\url{http://dx.doi.org/10.1371/journal.pone.0014735}
\item[2008]
\textbf{Anderson, S.C.}, H.K. Lotze, and N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic-level fisheries in Atlantic
Canada. Can. J. Fish. Aquat. Sci. 65(12): 2553--2571.
\url{http://dx.doi.org/10.1139/F08-156}.
(\href{papers/Anderson_etal_2008_knowledge.pdf}{PDF}) Eighth most-read
CJFAS article in 2008.
\end{description}

\section{Manuscripts in review}

\begin{description}
\item[2013]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Ecological prophets:
Quantifying metapopulation portfolio effects. Completing minor revisions
for Methods in Ecology and Evolution.
\item[2013]
O'Regan, S.M., W.J. Palen, \textbf{S.C. Anderson}. Climate warming
mediates negative impacts of rapid pond drying for three amphibian
species. In review at Ecology.
\item[2013]
Artelle K.A., \textbf{S.C. Anderson}, A.B. Cooper, Paul C. Paquet, J.D.
Reynolds, C.T. Darimont. Confronting uncertainty in wildlife management:
performance of grizzly bear management in British Columbia, Canada. In
review at PLOS ONE.
\end{description}
